\vspace{1.0cm} \LARGE \bf Revealing the competitive dynamics of influenza viruses through evolutionary cartography \\ or \\ Charting antigenic drift and evolutionary success of seasonal influenza lineages

### 2013-01-22 flux.tex
Rewrite abstract.

\affil[1]{Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK}
\affil[2]{Department of Biomathematics, David Geffen School of Medicine at UCLA, University of California, Los Angeles CA, USA}
\affil[3]{Department of Human Genetics, David Geffen School of Medicine at UCLA, University of California, Los Angeles CA, USA}
\affil[4]{Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles CA, USA}
\affil[5]{Department of Microbiology and Immunology, Katholieke Universiteit Leuven, Leuven, Belgium}
\affil[6]{Division of Virology, MRC National Institute for Medical Research, Mill Hill, London, UK}
\affil[7]{Department of Zoology, University of Cambridge, Cambridge, UK.}
\affil[8]{Department of Virology, Erasmus Medical Centre, Rotterdam, Netherlands.}
\affil[9]{Fogarty International Center, National Institutes of Health, Bethesda, MD, USA.}

@@

Influenza viruses undergo continual antigenic evolution allowing mutant viruses to evade immunity acquired by the host population to previous virus strains.
However, the extent to which antigenic evolution determines competitive dynamics within and between virus clades A/H3N2, A/H1N1, B/Victoria and B/Yamagata has remained unclear.
Here, we develop a novel cartographic approach to simultaneously characterize the genetic and antigenic evolution of viral lineages.
Through this approach, we determine historical rates and patterns of antigenic drift in all four viral clades and we show that antigenic drift influences year-to-year variability in clade dominance.
We investigate the selective underpinnings for differing antigenic dynamics across clades and suggest that pleiotropic effects linked to antigenic mutation determine selective outcomes.

@@

%%% errortable %%%
\begin{table}[h]
	\centering
	\caption{\textbf{Average absolute prediction error of log$_2$ HI titer for test data across models and datasets.}}
	\label{errortable}	
	\makebox[\textwidth][c]{
	\begin{tabular}{ c c c c c c c c } 
	\hline
	\multicolumn{4}{c}{} 	&	\multicolumn{4}{c}{Test error}  \\
	Model	&	Dimen 	& 	Serum/virus effects 	& Location Prior					& 	A/H3N2	& 	A/H1N1	& 	B/Vic	& 	B/Yam	\\
	\hline		
	1		&	2D 		&	Fixed $\ses$ 			& Uninformed						&	0.93 	&	0.78 	&	0.80	&	0.93	\\
	2		&	2D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.91 	&	0.74 	&	0.77	&	0.83	\\	
	3		&	1D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	1.07	&	0.92	&	0.94	&	1.08	\\		
	4		&	3D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.91	&	0.74 	&	0.79	&	0.84	\\
	5		&	4D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.95	&	0.75 	&	0.81	&	0.86	\\
	6		&	5D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.99	&	0.76 	&	0.84	&	0.88	\\				
	7		&	2D 		&	Estimated $\ses, \ves$	& $\beta,\virussd,\serumsd$			&	0.77	&	0.71 	&	0.65	&	0.73	\\	
	8		&	2D 		&	Estimated $\ses, \ves$	& $\beta,\virussd,\serumsd,\tree$	&	0.76	&	0.71 	&	0.64	&	0.72	\\		
	\hline
	\end{tabular}	
	}
\end{table}

>>>>

\affil[1]{Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK}
\affil[2]{Department of Biomathematics, David Geffen School of Medicine at UCLA, University of California, Los Angeles CA, USA}
\affil[3]{Department of Human Genetics, David Geffen School of Medicine at UCLA, University of California, Los Angeles CA, USA}
\affil[4]{Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles CA, USA}
\affil[5]{Department of Microbiology and Immunology, Katholieke Universiteit Leuven, Leuven, Belgium}
\affil[6]{Division of Virology, MRC National Institute for Medical Research, Mill Hill, London, UK}
\affil[7]{Department of Zoology, University of Cambridge, Cambridge, UK}
\affil[8]{Department of Virology, Erasmus Medical Centre, Rotterdam, Netherlands}
\affil[9]{Fogarty International Center, National Institutes of Health, Bethesda, MD, USA}

@@

Influenza viruses undergo continual antigenic evolution allowing mutant viruses to evade immunity acquired by the host population to previous virus strains.
Antigenic phenotype is often assessed through pairwise measurement of cross-reactivity between influenza strains using the hemagglutination inhibition (HI) assay.
Here, we extend previous approaches to antigenic cartography, which seeks to place strains on an antigenic map, such that distances on this map best recapitulate titers observed across multiple HI assays.
In our model, we simultaneously characterize antigenic and genetic evolution by including an evolutionary model in which antigenic location diffuses over a shared virus phylogeny.
Using HI data for influenza lineages A/H3N2, A/H1N1, B/Victoria and B/Yamagata, we determine average rates of antigenic drift for each clade, as well as year-to-year variability in the rate of drift.
Through comparison with epidemiological data, we show that antigenic drift influences overall incidence across clades, and demonstrate a year-to-year correlation between drift and incidence.
We investigate the selective underpinnings for differing antigenic dynamics across clades and show that A/H3N2 benefits from both a higher influx of new antigenic mutations and also from more efficient conversion of antigenic variation into fixed differences relative to other influenza clades.
This work does much to elucidate the competitive dynamics of influenza lineages, but also makes possible substantial future advances in influenza and other antigenically-variable pathogens by providing a model that intimately combines molecular and antigenic evolution.

@@

%%% errortable %%%
\begin{table}[h]
	\centering
	\caption{\textbf{Average absolute prediction error of log$_2$ HI titer for test data across models and datasets.}}
	\label{errortable}	
	\makebox[\textwidth][c]{
	\begin{tabular}{ c c c c c c c c } 
	\hline
	\multicolumn{4}{c}{} 	&	\multicolumn{4}{c}{Test error}  \\
	Model	&	Dimen 	& 	Serum/virus effects 	& Location Prior					& 	A/H3N2	& 	A/H1N1	& 	B/Vic	& 	B/Yam	\\
	\hline		
	1		&	2D 		&	Fixed $\ses$ 			& Diffuse							&	0.93 	&	0.78 	&	0.80	&	0.93	\\
	2		&	2D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.91 	&	0.74 	&	0.77	&	0.83	\\	
	3		&	1D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	1.07	&	0.92	&	0.94	&	1.08	\\		
	4		&	3D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.91	&	0.74 	&	0.79	&	0.84	\\
	5		&	4D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.95	&	0.75 	&	0.81	&	0.86	\\
	6		&	5D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.99	&	0.76 	&	0.84	&	0.88	\\				
	7		&	2D 		&	Estimated $\ses, \ves$	& $\beta,\virussd,\serumsd$			&	0.77	&	0.71 	&	0.65	&	0.73	\\	
	8		&	2D 		&	Estimated $\ses, \ves$	& $\beta,\virussd,\serumsd,\tree$	&	0.76	&	0.71 	&	0.64	&	0.72	\\		
	\hline
	\end{tabular}	
	}
\end{table}
